---
title: "SDS 315 Homework 2"
output:
  pdf_document: default
  html_document: default
  word_document: default
date: "2025-01-28"
---

Github Repository Link with Rmd Code File - https://github.com/vinaenae/SDS-315-Homework-2.git


```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, echo = FALSE, message = FALSE)
library(readr)
library(dplyr)
library(stringr)
library(ggplot2)
library(lubridate)
library(knitr)

```

```{r, echo = FALSE}

profs <- read_csv("SDS 315/profs.csv", show_col_types = FALSE)

bikeshare <- read_csv("SDS 315/bikeshare.csv", show_col_types = FALSE)

capmetro <- read_csv("SDS 315/capmetro.csv", show_col_types = FALSE)

billboard <- read_csv("SDS 315/billboard.csv", show_col_types = FALSE)
```


## Problem 1
# Part A
```{r, echo = FALSE}
# Problem 1: Beauty, or not, in the classroom
ggplot(profs, aes(x = eval)) +
  geom_histogram(binwidth = 0.2, fill = "blue", color = "black") +
  labs(title = "Distribution of Course Evaluation Scores", x = "Evaluation Score", y = "Frequency") +
  theme_minimal()
```
Through the histogram of the distribution of course evaluations, we can see that the relationship between the evaluation score and frequency is skewed left. We can see that there there is as maximum frequency of 60+ around an evaluation score of 4, and a minimum at the Evaluation Score of around 2.3 that corresponds to around 1 for the frequency. 


# Part B
```{r, echo = FALSE}
ggplot(profs, aes(x = native, y = eval, fill = native)) +
  geom_boxplot() +
  labs(title = "Course Evaluations by Native English Speaker Status", x = "Native English Speaker", y = "Evaluation Score") +
  theme_minimal()
```
For this boxplot, we can see that in general there is a greater evaluation score corresponding to Professor that is a Native English Speaker as compared to a non-native English Speaker. This is because the Native Speaker has a median of around 4 while the Non-Native Speaker has a median of around 3.6. For the Native Speaker though, we can see that there are some outliers with a low evaluation score.


# Part C
```{r, echo = FALSE}
ggplot(profs, aes(x = eval)) +
  geom_histogram(binwidth = 0.2, fill = "purple", color = "black") +
  facet_wrap(~gender, nrow = 2) +
  labs(title = "Evaluation Scores by Gender", x = "Evaluation Score", y = "Frequency") +
  theme_minimal()
```
Through the histogram, we can see that there is a higher evaluation score for males in general with more male teachers having a higher evaluation score and higher frequency. The male evaluation score distribution is skewed left while the female evaluation score distribution seems to be normally distributed. There is also a higher max frequency on the male side with around 40 plus with a evaluation score of around 4.3.


# Part D
```{r, echo = FALSE}
ggplot(profs, aes(x = beauty, y = eval)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", color = "red") +
  labs(title = "Relationship Between Attractiveness and Evaluation Scores", x = "Attractiveness", y = "Evaluation Score") +
  theme_minimal()

cor(profs$beauty, profs$eval)
```
Through this scatter plot of data between attractiveness and evaluation scores, we can see through the regression line that there is a moderately positive, linear relationship between the two variables. Through the correlation method, we can see that 0.189, which is pretty close to 0 but positive, defines this moderate relationship.

## Problem 2
# Part A
```{r, echo = FALSE}
# Problem 2: Bike Sharing

avg_hourly <- bikeshare %>%
  group_by(hr) %>%
  summarize(avg_total = mean(total))

ggplot(avg_hourly, aes(x = hr, y = avg_total)) +
  geom_line(color = "blue") +
  labs(title = "Average Hourly Bike Rentals", x = "Hour of the Day", y = "Average Rentals") +
  theme_minimal()
```
Through this line plot, we can clearly see the timings throughout the day that affects average rentals for bikes. Usually, around the morning (5 - 10 hours into the day) there is a spike in rentals and there is another spike closer to the end of the day around 17 hours. 

# Part B
```{r, echo = FALSE}
avg_by_workingday <- bikeshare %>%
  group_by(hr, workingday) %>%
  summarize(avg_total = mean(total))

ggplot(avg_by_workingday, aes(x = hr, y = avg_total, color = factor(workingday))) +
  geom_line() +
  facet_wrap(~workingday) +
  labs(title = "Average Hourly Rentals by Working Day", x = "Hour", y = "Average Rentals") +
  theme_minimal()
```
According to these side-by-side line plots, we can see that on working days there are a significantly more amount of average rentals overall. These plots also vary in the sense that they have different hours where there are max rentals. For example, for the non-working days, average rentals reach the peak when hours are around 13. On the contrary, on the working days, the average rentals reach their peak when hours around 7 and hours are around 17.


# Part C
```{r, echo = FALSE}
ridership_9am <- bikeshare %>%
  filter(hr == 9) %>%
  group_by(weathersit, workingday) %>%
  summarize(avg_total = mean(total))

ggplot(ridership_9am, aes(x = factor(weathersit), y = avg_total, fill = factor(workingday))) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_wrap(~workingday) +
  labs(title = "9 AM Ridership by Weather and Working Day", x = "Weather Situation", y = "Average Rentals") +
  theme_minimal()
```
According to the bar plots that display the variables of working age, weather situation, and average rentals, we can see many different insights. On the right plot, there are the max amount of rentals in weather conditions of clear, few clouds, partly cloudy alongside possible mist, and is a working day. Again, like the last plot, there are more average rentals during working days. On the left plot, we can see that there are the highest amount of rentals during cloud or clear days which does match with the right plot.

## Problem 3
# Part A
```{r, echo = FALSE}
# Problem 3: Capital Metro UT Ridership
capmetro <- capmetro %>%
  mutate(
    hour_of_day = hour(timestamp),
    day_of_week = wday(timestamp, label = TRUE)
  )

# Group and summarize the data
boarding_hourly <- capmetro %>%
  group_by(hour_of_day, day_of_week) %>%
  summarize(avg_boarding = mean(boarding, na.rm = TRUE))

# Plot the data
ggplot(boarding_hourly, aes(x = hour_of_day, y = avg_boarding, color = day_of_week)) +
  geom_line() +
  labs(
    title = "Boarding Activity by Hour and Day",
    x = "Hour of Day",
    y = "Average Boardings"
  ) +
  theme_minimal()
```
For the line graph, we can see the variables of hour of day, day of week, and average rentals accordingly. We can see that there are the highest amount of average rentals when the hour of days is around 15-20. More specifically, Saturdays and Sunday obviously have the lowest average rentals as indicated by the purple and yellow colors, and the working days all have the highest rentals.

# Part B
```{r, echo = FALSE}
capmetro <- capmetro %>%
  mutate(
    weekend = ifelse(wday(timestamp, label = TRUE) %in% c("Sat", "Sun"), "Weekend", "Weekday")
  )

ggplot(capmetro, aes(x = temperature, y = boarding)) +
  geom_point(alpha = 0.5) +
  facet_wrap(~weekend) +
  labs(
    title = "Boarding Activity vs Temperature",
    x = "Temperature (°F)",
    y = "Boarding"
  ) +
  theme_minimal()
```
According to the scatterplot of data describing boarding activity and temperature, we can see that working days where people commute a lot to work have a higher overall boarding rate. On the contrary, weekends have way less. The lower temperatures less than 40 degrees and hot temperatures do affect travel to some extent but not much. ON average, on both the plots, there arew more boarding rates when temperatures are between 45 and 90. 

## Problem 4
# Part A
```{r, echo = FALSE}
# Problem 4: Wrangling the Billboard Top 100
top_songs <- billboard %>%
  group_by(performer, song) %>%
  summarise(count = n()) %>%
  arrange(desc(count)) %>%
  head(10)

kable(top_songs, col.names = c("Performer", "Song", "Times Charted"), align = "c")
```
According to the table displaying th performers, songs, and count, we can see from the pattern of the max count to the min count that Imagine Dragons are the best performers and Carrie Underwood is the lowest performer of the dataset.

# Part B
```{r, echo = FALSE}
filtered_data <- billboard %>%
  filter(year > 1958 & year < 2021)

# Count the number of unique songs per year
musical_diversity <- filtered_data %>%
  group_by(year) %>%
  summarise(unique_songs = n_distinct(song))

ggplot(musical_diversity, aes(x = year, y = unique_songs)) +
  geom_line(color = "blue") +
  labs(
    title = "Musical Diversity of the Billboard Top 100 Over Time",
    x = "Year",
    y = "Number of Unique Songs",
    caption = "Excludes data for 1958 and 2021 due to incomplete data"
  ) +
  theme_minimal()
```
According to the line plot, we can see that there is a diverse amount of unique songs for every group of years. Earlier, around 1970, there was a huge spioke in number of unique songs. After this, the number slowly started declining close to the 2000 era, and again ramping up during the 21st century.

# Part C
```{r, echo = FALSE}
ten_week_hits <- billboard %>%
  group_by(performer, song) %>%
  summarise(total_weeks = n()) %>%
  filter(total_weeks >= 10) %>%
  ungroup()

# Count the number of ten-week hits for each artist
artist_hits <- ten_week_hits %>%
  group_by(performer) %>%
  summarise(ten_week_hits = n()) %>%
  filter(ten_week_hits >= 30) %>%
  arrange(desc(ten_week_hits))

# Plot the bar chart
ggplot(artist_hits, aes(x = reorder(performer, -ten_week_hits), y = ten_week_hits)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  coord_flip() +
  labs(
    title = "Artists with 30 or More Ten-Week Hits",
    x = "Artist",
    y = "Number of Ten-Week Hits",
    caption = "Data from Billboard Top 100 (1958-2021)"
  ) +
  theme_minimal()
```
According to this detailed bar plot that depicts an increasing number of ten week hits for every artist, we can see that Neil Diamond has the least amount of ten-week hits and Elton John had the most amount of hits. Also, specifically to the code, we had to flip the x and y axis to represent them clearer.